---
title: "HW 3"
author: "Tallapaneni Venkateshwara Chowdary"
date: "2024-10-28"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r}
library(tidyverse)
library(ggplot2)
library(ggtext)
library(tidyr)
library(ggpubr)
library(jpeg)
library(scales)

# read the data
income <- read.csv("data/income.csv")

```

Question 1's Answer:
```{r}
income1 <- income |> 
  unite(Class, c(Class, Average_Income), sep = " | ", remove = FALSE) %>%
  # reordering the y-axis values
  mutate(Class = fct_relevel(Class, 
            "$1000 AND OVER | 1125", "$750-1000 | 880", "$500-750 | 547", 
            "$400-500 | 433.82", "$300-400 | 335.66", "$200-300 | 249.45", 
            "$100-200 | 139.1")) %>%
  select(-Average_Income) %>%
  pivot_longer(
    cols = !Class,
    names_to = "Expense_Type", 
    values_to = "Values"
  )

# Reording expense types to be the same as in the picture 
income1$ANNUAL_EXPENDITURE_FOR <- factor(income1$Expense_Type, levels = c('Rent', 'Food', 'Clothes', 'Tax', 'Other'),ordered = TRUE)

# Calculating the Percentage of each expense
income1 <- income1 %>%
  group_by(Class) %>%
  mutate(Percentage = Values / sum(Values) * 100) %>%
  ungroup()

img <- readJPEG("paper-3653357_1920.jpg")

  income1 |>
  ggplot(aes(fill=ANNUAL_EXPENDITURE_FOR, y=Class, x=Values)) + 
   background_image(img)+
    geom_bar(stat="identity", position = position_fill(reverse = TRUE)) +
    geom_text(aes(label = paste0(round(Percentage, 1), "%")), 
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "grey90", size = 3, 
            data = income1) +
  theme(legend.position = "top",
        # I couldn't get the background as image so I imitated the color
        plot.background = element_rect(fill = "#e8c079"),
        legend.background = element_rect(fill = "#e8c079")) +
  scale_fill_manual(values=c("#141718", "#7a557c", "#de9f8c" , "#a59f9b" , "#cab9a7")) +
  labs(
    title = "INCOME AND EXPENDITURE OF 150 NEGRO FAMILIES IN ATLANTA,GA.,U.S.A", 
    fill = "ANNUAL EXPENDITURE FOR ", 
    y = 'Class | ACTUAL AVERAGE',
    x = "FOR FURTHER STATISTICS RAISE THIS FRAME/Expense Percentage "
  )
```

Question 2's Answer:
  From the plot, I can safely say that absolutely no one regardless of race, Gender, Race, Ethnicity, and Profession trust the information about the vaccines due to the low response scores. We can also say that Native Hawaiians and Asians are the same in regard to their view about recommending vaccine to others - a strong 'NO'. Finally the information that shocked me the most that 21-25 and 26-30 include the people who are least confident in the scientific vetting process of the vaccines because I assumed these age groups have the most literates - Undergraduates and Graduates and they would have the most confidence in the vetting process.


Question 3's Answer:
```{r}
# Cleaning up the csv data
survey_data <- read.csv("data/covid-survey.csv")
colnames(survey_data) <- survey_data[1, ]
survey_data <- survey_data[-1, ]
rownames(survey_data) <- NULL
head(survey_data)
```

```{r}
survey_data <- survey_data %>%
  mutate(
    across(starts_with("resp_"), as.numeric),
    exp_age_bin = as.numeric(exp_age_bin),
    exp_gender = as.factor(exp_gender),
    exp_race = as.factor(exp_race),
    exp_ethnicity = as.factor(exp_ethnicity),
    exp_profession = as.factor(exp_profession),
    exp_flu_vax = as.factor(exp_flu_vax),
    exp_already_vax = as.factor(exp_already_vax)
  )
```

```{r}
# Create bins for explanatory variables with corresponding labels
survey_data <- survey_data %>%
  mutate(
    exp_age_bin = factor(exp_age_bin, levels = c("0", "20", "25", "30"), 
                         labels = c("<20", "21-25", "26-30", ">30")),
    
    exp_gender = factor(exp_gender, levels = c("0", "1", "3", "4"),  
                        labels = c("Male", "Female", "Non-binary / third gender", "Prefer not to say")),
  
    exp_race = factor(exp_race, levels = c("1", "2", "3", "4", "5"), 
                      labels = c("American Indian/Alaska Native", "Asian", 
                                 "Black/African American", "Native Hawaiian/Other Pacific Islander", "White")),

    exp_ethnicity = factor(exp_ethnicity, levels = c("1", "2"), 
                           labels = c("Hispanic/Latino", "Non-Hispanic/Non-Latino")),

    exp_profession = factor(exp_profession, levels = c("0", "1"), 
                            labels = c("Medical", "Nursing")),

    exp_already_vax = factor(exp_already_vax, levels = c("0", "1"), 
                             labels = c("No", "Yes")),

    exp_flu_vax = factor(exp_flu_vax, levels = c("0", "1"), 
                         labels = c("No", "Yes"))
  )
```


```{r}
# Pivot the data longer twice
covid_survey_longer <- survey_data %>%
  pivot_longer(
    cols = starts_with("exp_"),
    names_to = "explanatory",
    values_to = "explanatory_value"
  ) %>%
  filter(!is.na(explanatory_value)) %>%
  pivot_longer(
    cols = starts_with("resp_"),
    names_to = "response",
    values_to = "response_value"
  )
```


```{r}
# Calculate summary statistics by grouping by each explanatory and response variable
covid_survey_summary_stats_by_group <- covid_survey_longer %>%
  group_by(explanatory, explanatory_value, response) %>%
  summarize(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low = quantile(as.numeric(response_value), 0.10, na.rm = TRUE),
    high = quantile(as.numeric(response_value), 0.90, na.rm = TRUE)
  )
```


```{r}
# Calculate summary statistics without conditioning on explanatory variables
covid_survey_summary_stats_all <- covid_survey_longer %>%
  group_by(response) %>%
  summarize(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low = quantile(as.numeric(response_value), 0.10, na.rm = TRUE),
    high = quantile(as.numeric(response_value), 0.90, na.rm = TRUE),
    explanatory = "All",
    explanatory_value = ""
  )
```


```{r}
# Combine both sets of summary statistics
covid_survey_summary_stats <- bind_rows(
  covid_survey_summary_stats_all,
  covid_survey_summary_stats_by_group
)
```


```{r}
# Update the mapping for response labels with full texts in a smaller font size
response_labels <- c(
  "resp_safety" = "Based on my understanding, I believe the vaccine is safe.",
  "resp_feel_safe_at_work" = "Getting the vaccine will make me feel safer at work",
  "resp_concern_safety" = "I am concerned about the safety and side effects of the vaccine",
  "resp_confidence_science" = "I am confident in the scientific vetting process for the new COVID vaccines.",
  "resp_trust_info" = "I trust the information that I have received about the COVID-19 vaccines.",
  "resp_will_recommend" = "I will recommend the vaccine to family, friends, and community members"
)
```


```{r}
# Define the order for the response and explanatory variables in covid_survey_summary_stats
covid_survey_summary_stats <- covid_survey_summary_stats %>%
  mutate(
    response = factor(response, levels = c(
      "resp_safety",  # Based on my understanding, I believe the vaccine is safe.
      "resp_feel_safe_at_work",  # Getting the vaccine will make me feel safer at work
      "resp_concern_safety",  # I am concerned about the safety and side effects of the vaccine
      "resp_confidence_science",  # I am confident in the scientific vetting process for the new COVID vaccines.
      "resp_trust_info",  # I trust the information that I have received about the COVID-19 vaccines.
      "resp_will_recommend"  # I will recommend the vaccine to family, friends, and community members
    )),
    explanatory = factor(explanatory, levels = c(
      "All", "exp_age_bin", "exp_gender", "exp_race", "exp_ethnicity", 
      "exp_profession", "exp_already_vax", "exp_flu_vax"
    ))
  )
```

```{r}
# Define the mapping for explanatory labels
explanatory_labels <- c(
  "exp_age_bin" = "Age",
  "exp_gender" = "Gender",
  "exp_race" = "Race",
  "exp_ethnicity" = "Ethnicity",
  "exp_profession" = "Profession",
  "exp_already_vax" = "Had COVID vaccine",
  "exp_flu_vax" = "Had flu vaccine this year"
)
```

```{r fig.width=14, fig.height=9}
# Plot the updated data with full text labels
ggplot(covid_survey_summary_stats, aes(x = mean, y = explanatory_value, color = explanatory_value)) +
  geom_point(color = "black") +
  geom_errorbarh(aes(xmin = low, xmax = high), height = 0.2, color = "black") +
  facet_grid(
    explanatory ~ response,
    scales = "free_y",
    labeller = labeller(
      response = as_labeller(response_labels, default = label_wrap_gen(width = 35)), 
      explanatory = as_labeller(explanatory_labels)
    )
  ) +
  theme_minimal() +
  labs(
    title = "Survey Responses on COVID-19 Vaccines",
    subtitle = "Mean Likert Score with 10th to 90th Percentile Error Bars",
    x = "Mean Likert Score",
    y = NULL
  ) +
  theme(
    strip.background = element_rect(fill = "gray90"),
    strip.text.y = element_text(angle = 0),
    strip.text.x = element_text(size = 8),  # Smaller font size for facet titles
    legend.position = "none"
  )
```

Question 4's Answer:
```{r fig.width=14, fig.height=9}
# Calculate summary statistics by grouping by each explanatory and response variable
covid_survey_summary_stats_by_group_quartiles <- covid_survey_longer %>%
  group_by(explanatory, explanatory_value, response) %>%
  summarize(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low = quantile(as.numeric(response_value), 0.25, na.rm = TRUE),  # 25th percentile
    high = quantile(as.numeric(response_value), 0.75, na.rm = TRUE)  # 75th percentile
  )
```

```{r}
# Calculate summary statistics without conditioning on explanatory variables
covid_survey_summary_stats_all_quartiles <- covid_survey_longer %>%
  group_by(response) %>%
  summarize(
    mean = mean(as.numeric(response_value), na.rm = TRUE),
    low = quantile(as.numeric(response_value), 0.25, na.rm = TRUE),  # 25th percentile
    high = quantile(as.numeric(response_value), 0.75, na.rm = TRUE),  # 75th percentile
    explanatory = "All",
    explanatory_value = ""
  )
```

```{r}
# Combine both sets of summary statistics
covid_survey_summary_stats_quartiles <- bind_rows(
  covid_survey_summary_stats_all_quartiles,
  covid_survey_summary_stats_by_group_quartiles
)
```

```{r}
# Define the order for the response and explanatory variables in covid_survey_summary_stats_quartiles
covid_survey_summary_stats_quartiles <- covid_survey_summary_stats_quartiles %>%
  mutate(
    response = factor(response, levels = c(
      "resp_safety", 
      "resp_feel_safe_at_work",  
      "resp_concern_safety",  
      "resp_confidence_science",  
      "resp_trust_info",  
      "resp_will_recommend"  
    )),
    explanatory = factor(explanatory, levels = c(
      "All", "exp_age_bin", "exp_gender", "exp_race", "exp_ethnicity", 
      "exp_profession", "exp_already_vax", "exp_flu_vax"
    ))
  )
```

```{r}
# Plot the updated data with 25th and 75th percentiles
ggplot(covid_survey_summary_stats_quartiles, aes(x = mean, y = explanatory_value, color = explanatory_value)) +
  geom_point(color = "black") +
  geom_errorbarh(aes(xmin = low, xmax = high), height = 0.2, color = "black") +
  facet_grid(
    explanatory ~ response,
    scales = "free_y",
    labeller = labeller(
      response = as_labeller(response_labels, default = label_wrap_gen(width = 35)), 
      explanatory = as_labeller(explanatory_labels)
    )
  ) +
  theme_minimal() +
  labs(
    title = "Survey Responses on COVID-19 Vaccines (Quartiles)",
    subtitle = "Mean Likert Score with 25th to 75th Percentile Error Bars",
    x = "Mean Likert Score",
    y = NULL
  ) +
  theme(
    strip.background = element_rect(fill = "gray90"),
    strip.text.y = element_text(angle = 0),
    strip.text.x = element_text(size = 8),
    legend.position = "none"
  )
```
The only major difference i could find was the removal of distribution lines for some plots due the reduction in percentile range. And as none of my interpretations were based on the distribution line I will say all my previous interpretations stay unchanged.

Question 5's Answer:
```{r}
survey_data <- survey_data %>%
  mutate(across(everything(), ~ as.character(.))) %>%  # Convert factors to characters
  mutate(across(everything(), ~ na_if(., ""))) %>%     # Replace empty strings with NA
  mutate(across(everything(), ~ as.factor(.)))          # Convert back to factors if needed

survey_data_clean <- survey_data |>
  select(contains("resp_")) |>
  pivot_longer(
    cols = everything(),
    names_to = "question", 
    values_to = "response",
  ) |>
  mutate(question = str_remove(question, "resp_")) |>
  drop_na(response)
```

```{r}
unique(survey_data_clean$response)
```


```{r}
survey_data_summary <- survey_data_clean %>% 
  group_by(question, response) %>% 
  count(name = "n_responses") %>% 
  group_by(question) %>% 
  mutate(percent_answers = n_responses / sum(n_responses)) %>% 
  ungroup() %>% 
  mutate(percent_answers_label = percent(percent_answers, accuracy = 1))
survey_data_summary
```

```{r}
survey_data_summary_diverging <- survey_data_summary %>%
  mutate(percent_answers = if_else(response %in% c("3", "4", "5"), percent_answers, -percent_answers)) %>% 
  mutate(percent_answers_label = percent(percent_answers, accuracy = 1))
survey_data_summary_diverging
```

```{r}
survey_data_summary_diverging_good_labels <- survey_data_summary_diverging %>%
  mutate(percent_answers_label = abs(percent_answers)) %>% 
  mutate(percent_answers_label = percent(percent_answers_label, accuracy = 1))
survey_data_summary_diverging_good_labels
```


```{r}
survey_data_summary_diverging_good_labels %>% 
  ggplot(aes(x = question, 
             y = percent_answers,
             fill = response)) +
  geom_col() +
  geom_text(aes(label = percent_answers_label),
            position = position_stack(vjust = 0.5),
            color = "white",
            fontface = "bold") +
  coord_flip() +
  scale_x_discrete() +
  scale_fill_viridis_d() +
  labs(title = "Diverging Bar Chart of the Survey Questions",
       x = NULL,
       fill = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "top")
```


```{r}
survey_data_summary_diverging_right_order <- survey_data_summary_diverging_good_labels  %>% 
  mutate(response = fct_relevel(response,
                               "2", "1", "3", "4", "5"),
         response = fct_rev(response)) 
survey_data_summary_diverging_right_order
```


```{r}
survey_data_summary_diverging_right_order %>%
  ggplot(aes(x = question, 
             y = percent_answers,
             fill = response)) +
  geom_col() +
  geom_text(
    aes(label = percent_answers_label),
    position = position_stack(vjust = 0.5),
    color = "white",
    fontface = "bold"
  ) +
  coord_flip() +
  scale_x_discrete() +
  scale_fill_viridis_d() +
  labs(title = "Diverging Bar Chart of the Survey Questions",
       x = NULL,
       fill = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "top")
```


```{r}
survey_data_summary_diverging_right_order %>%
  ggplot(aes(x = question, 
             y = percent_answers,
             fill = response)) +
  geom_col() +
  geom_text(aes(label = percent_answers_label),
            position = position_stack(vjust = 0.5),
            color = "white",
            fontface = "bold") +
  coord_flip() +
  scale_x_discrete() +
  scale_fill_viridis_d(breaks = c("1", "2", "3", "4", "5")) +
  labs(title = "Diverging Bar Chart of the Survey Questions",
       x = NULL,
       fill = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "top")
```

```{r}
survey_data_summary_diverging_right_order %>%
  ggplot(aes(x = question, 
             y = percent_answers,
             fill = response)) +
  geom_col() +
  geom_text(aes(label = percent_answers_label),
            position = position_stack(vjust = 0.5),
            color = "white",
            fontface = "bold") +
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(breaks = c("1", "2", "3", "4","5"),
                    values = c(
                      "1" = "darkorange3",
                      "2" = "orange",
                      "4" = "deepskyblue",
                      "5" = "deepskyblue4",
                      "3" = "#f7e0ad"
                    )) +
  labs(title = "Diverging Bar Chart of the Survey Questions",
       x = NULL,
       fill = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "top")
```

```{r}
survey_data_summary |>
ggplot(aes(x = question, y = percent_answers, fill = response)) +
  geom_bar(stat="identity", position = position_fill(reverse = TRUE)) +
  geom_text(aes(label = percent_answers_label), 
            position = position_fill(vjust = 0.5, reverse = TRUE)) +
  coord_flip() +
  scale_x_discrete() +
  # https://icolorpalette.com/download/palette/578493_color_palette.jpg
  scale_fill_manual(breaks = c("1", "2", "3", "4","5"),
                    values = c(
                      "1" = "darkorange3",
                      "2" = "orange",
                      "3" = "#f7e0ad",
                      "4" = "deepskyblue",
                      "5" = "deepskyblue4"
                    )) +
  labs(title = "100% Bar Chart of the Survey Questions",
       x = NULL,
       fill = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "top")
```

  The 100% Bar Chart is not very intuitive for this likert visualization. So I prefer using the Divergent Bar Chart instead for this likert visualization because I can have the freedom of selecting the "good" side of the responses like in this case I have chosen 3 and above(my good side) for one side and the rest(not good) for the other side.
  
References: 
  https://rfortherestofus.com/2021/10/diverging-bar-chart